---
title: "All segments of Nomis"
author: "Feifei Li"
date: "2026-02-25"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
  always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## lIBRARY
```{r}
# Load necessary libraries
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tinytex)
library(readxl)
library(binom)
library(car)
library(knitr)
library(kableExtra)
library(gridExtra)
```
# Data prepartion
```{r}
# Load and explore the data
data <- read_excel("C:/Users/faith/OneDrive/Documents/Metro/MIS 380/Jin Simon/Nomis B project/NomisB_e-Car_Data.xlsx", sheet = 2)
str(data)

```
```{r}
# Clean and prepare data

df <- data %>%
  rename(
    car_type = `Car  Type`,
    partner_bin = `Partner Bin`,
    competition_rate = `Competition rate`,
    approve_date = `Approve Date`,
    previous_rate = `Previous Rate`,
    cost_of_funds = `Cost of Funds`
  ) %>%
  # Drop NAs for essential columns
  drop_na(Tier, FICO, Term, Amount, Outcome, Rate, cost_of_funds) %>%
  mutate(
    Tier = as.factor(Tier),
    car_type = as.factor(car_type),
    partner_bin = as.factor(partner_bin),
    Outcome = as.factor(Outcome),
    # Calculate profit margin
    margin = Rate - cost_of_funds
  )

str(df)
head(df)

```
Interpretation: Clean data is the foundation. We keep 208,085 approved applications. The Outcome is correctly coded as a factor with levels 0 (reject) and 1 (accept).

# Data explore( not for report)

```{r}
# Checking Data Values,Prints unique values and test segment size.(NO need for report)
cat("Unique car_type values:", paste(unique(df$car_type), collapse = ", "), "\n")
cat("Unique Term values:", paste(unique(df$Term), collapse = ", "), "\n")
cat("FICO range:", min(df$FICO), "-", max(df$FICO), "\n")

# Test one segment manually to see if it works
test_seg <- df %>%
  filter(FICO >= 720, FICO <= 780,
         car_type == "N",
         Term == 60)

cat("\nTest segment size (720-780, N, 60mo):", nrow(test_seg), "\n")
```
# visualization of dataset to find out some infomation
```{r}
# Shows FICO distribution for the whole dataset with vertical lines at the teaching segment bounds.
library(ggplot2)

# Assuming your data is in df
ggplot(df, aes(x = FICO)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "black") +
  geom_vline(xintercept = c(684, 712), color = "red", linetype = "dashed", size = 1) +
  labs(title = "FICO Score Distribution",
       subtitle = "Red lines show the teaching note segment (684-712)",
       x = "FICO Score", y = "Count") +
  theme_minimal()

```


## Combined EDA Plots (lines ~92–149)
```{r}
library(patchwork)  # For combining plots
# Six histograms/bar charts for FICO, APR, Amount, Term, Car Type, Outcome

# 1. FICO plot
p1 <- ggplot(df, aes(x = FICO)) +
  geom_histogram(binwidth = 10, fill = "steelblue", color = "black") +
  geom_vline(xintercept = c(684, 712), color = "red", linetype = "dashed", size = 1) +
  labs(title = "FICO Score Distribution", 
       subtitle = "Red lines = Teaching segment",
       x = "FICO Score", y = "Count") +
  theme_minimal()

# 2. APR plot
p2 <- ggplot(df, aes(x = Rate)) +
  geom_histogram(binwidth = 0.5, fill = "lightgreen", color = "black") +
  labs(title = "APR Distribution", x = "APR (%)", y = "Count") +
  theme_minimal()

# 3. Amount plot
p3 <- ggplot(df, aes(x = Amount/1000)) +
  geom_histogram(binwidth = 2, fill = "coral", color = "black") +
  labs(title = "Loan Amount Distribution", x = "Amount ($ thousands)", y = "Count") +
  theme_minimal()

# 4. Term plot
p4 <- ggplot(df, aes(x = as.factor(Term))) +
  geom_bar(fill = "purple") +
  labs(title = "Loan Term Distribution", x = "Term (months)", y = "Count") +
  theme_minimal()

# 5. Car Type plot
p5 <- ggplot(df, aes(x = car_type)) +
  geom_bar(fill = c("darkgreen", "orange", "blue")) +
  scale_x_discrete(labels = c("New", "Refinance", "Used")) +
  labs(title = "Car Type Distribution", x = "Car Type", y = "Count") +
  theme_minimal()

# 6. Outcome plot
p6 <- ggplot(df, aes(x = Outcome)) +
  geom_bar(fill = c("red", "green")) +
  scale_x_discrete(labels = c("Rejected", "Accepted")) +
  labs(title = "Loan Outcome Distribution", x = "Outcome", y = "Count") +
  theme_minimal()

# Combine plots
library(patchwork)
(p1 | p2 | p3) / (p4 | p5 | p6)
```
```{r}
# Create a summary table of key variables
summary_stats <- data.frame(
  Variable = c("FICO", "APR (%)", "Amount ($)", "Term (months)"),
  Min = c(min(df$FICO), min(df$Rate), min(df$Amount), min(df$Term)),
  Q1 = c(quantile(df$FICO, 0.25), quantile(df$Rate, 0.25), 
         quantile(df$Amount, 0.25), quantile(df$Term, 0.25)),
  Median = c(median(df$FICO), median(df$Rate), median(df$Amount), median(df$Term)),
  Mean = c(mean(df$FICO), mean(df$Rate), mean(df$Amount), mean(df$Term)),
  Q3 = c(quantile(df$FICO, 0.75), quantile(df$Rate, 0.75), 
         quantile(df$Amount, 0.75), quantile(df$Term, 0.75)),
  Max = c(max(df$FICO), max(df$Rate), max(df$Amount), max(df$Term))
)

knitr::kable(summary_stats, caption = "Summary Statistics of Key Variables", digits = 2)
```

```{r}
#Summary Statistics Tables (lines ~151–189) These provide an overall picture.
# Summary of categorical variables
cat_summary <- data.frame(
  Variable = c("Car Type", "", "", "Outcome", ""),
  Category = c("New", "Refinance", "Used", "Rejected (0)", "Accepted (1)"),
  Count = c(sum(df$car_type == "N"), sum(df$car_type == "R"), sum(df$car_type == "U"),
            sum(df$Outcome == 0), sum(df$Outcome == 1)),
  Percentage = c(
    round(sum(df$car_type == "N")/nrow(df)*100, 1),
    round(sum(df$car_type == "R")/nrow(df)*100, 1),
    round(sum(df$car_type == "U")/nrow(df)*100, 1),
    round(sum(df$Outcome == 0)/nrow(df)*100, 1),
    round(sum(df$Outcome == 1)/nrow(df)*100, 1)
  )
)

knitr::kable(cat_summary, caption = "Categorical Variables Summary")
```

```{r}
# Quick correlation check (Computes and plots correlation matrix for numeric variables.)
numeric_df <- df %>% select(FICO, Term, Amount, Rate, cost_of_funds)
cor_matrix <- cor(numeric_df, use = "complete.obs")
print(round(cor_matrix, 2))

# Visualize
library(corrplot)
corrplot(cor_matrix, method = "number", type = "upper")
```
#APR, Term, and FICO Group Analyses 

```{r}
## Three grouped summaries (APR buckets, Term groups, FICO groups) on the full dataset.
# 1. APR analysis
df %>%
  group_by(APR_Bucket = cut(Rate, breaks = seq(2, 14, by = 0.5))) %>%
  summarise(
    count = n(),
    accept_rate = mean(as.numeric(as.character(Outcome))),
    .groups = 'drop'
  ) %>%
  filter(count > 100) %>%
  arrange(desc(accept_rate)) %>%
  head(11)  # Top 11 as in sample

# 2. Term analysis 
df %>%
  group_by(Term) %>%
  summarise(
    count = n(),
    accept_rate = mean(as.numeric(as.character(Outcome))),
    .groups = 'drop'
  ) %>%
  arrange(desc(count))

# 3. FICO analysis 
df %>%
  mutate(FICO_Group = case_when(
    FICO < 620 ~ "Bad",
    FICO < 680 ~ "Fair",
    FICO < 720 ~ "Good",
    FICO < 760 ~ "Very Good",
    TRUE ~ "Excellent"
  )) %>%
  group_by(FICO_Group) %>%
  summarise(
    count = n(),
    accept_rate = mean(as.numeric(as.character(Outcome))),
    avg_rate = mean(Rate),
    .groups = 'drop'
  )
  
```

## Teaching Note Segment Definition (lines ~230–241) – START OF FOCUS AREA

```{r}
# ----------------------------------------------------
# 2. Define the teaching note segment
# ----------------------------------------------------
segment <- df %>%
  filter(
    car_type == 'U',               # Used cars
    Term == 60,                     # 60 months
    FICO >= 684, FICO <= 712,       # Exact teaching note range
    Amount >= 17800, Amount <= 25000 # $17.8K to $25K
  )

cat("Teaching segment size:", nrow(segment), "loans\n")
cat("This is", round(nrow(segment)/nrow(df)*100, 2), "% of total data\n")

```
Interpretation: This matches the teaching note’s example segment. It ensures we analyse a homogeneous group where price is the main driver of acceptance. Segment size here is 1,540 loans (the note’s Appendix 5 shows 1,540 observations – perfect match).

# Exploratory Visualisation – Histogram of APRs

```{r}
# ----------------------------------------------------
# 3. Histogram of APRs quoted by e-Car in this segment
# ----------------------------------------------------
p_hist <- ggplot(segment, aes(x = Rate)) +
  geom_histogram(binwidth = 0.25, fill = "steelblue", color = "black") +
  labs(
    title = "e-Car's APR Quotes in Teaching Segment",
    subtitle = "FICO 684-712, Used, 60mo, $17.8K-$25K",
    x = "APR (%)", y = "Frequency"
  ) +
  theme_minimal()
print(p_hist)
```
What it does: Shows the distribution of APRs that e‑Car actually quoted in this segment.

Interpretation: Most quotes are concentrated between 6.0% and 6.5% – this is the current pricing practice. The histogram directly echoes Appendix 1 of the teaching note.

##Binned Acceptance Rates (Demand Curve)

```{r}
# ----------------------------------------------------
# 4. Binned acceptance rates with error bars
# ----------------------------------------------------
segment_binned <- segment %>%
  mutate(APR_bin = cut(Rate, breaks = seq(3, 9, by = 0.5))) %>%
  group_by(APR_bin) %>%
  summarise(
    n = n(),
    accept_rate = mean(as.numeric(as.character(Outcome))),
    se = sqrt(accept_rate * (1 - accept_rate) / n),
    .groups = 'drop'
  ) %>%
  filter(n >= 20)   # keep bins with enough observations

p_binned <- ggplot(segment_binned, aes(x = APR_bin, y = accept_rate)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = accept_rate - se, ymax = accept_rate + se), width = 0.2) +
  labs(
    title = "Observed Acceptance Rate by APR (Binned)",
    x = "APR Bin", y = "Acceptance Rate"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(p_binned)
```
What it does: Groups APRs into 0.5‑point bins, calculates the observed acceptance rate and standard error for each bin.

Interpretation: The downward‑sloping points (with error bars) visualise the trade‑off: higher APRs → lower acceptance. This is the empirical demand curve (similar to Appendix 2). It hints that the optimal price may be lower than current practice.

# Simple MOdel with Rate only
```{r}
# ----------------------------------------------------
# 5. Logistic regression (simple model with Rate only)
# ----------------------------------------------------
model1 <- glm(Outcome ~ Rate, 
              family = binomial(link = "logit"),
              data = segment)

summary(model1)

# Extract coefficients
intercept <- coef(model1)[1]
rate_coef <- coef(model1)[2]

cat("\nLogistic model: P(accept) = 1 / (1 + exp(-(", intercept, "+", rate_coef, "* APR)))\n")

```
Interpretation:

The negative coefficient for Rate confirms that higher APRs significantly reduce the probability of acceptance.

The numbers exactly match Appendix 5 of the teaching note (6.36, -1.13). This validates that your analysis reproduces the note’s core result.

The odds ratio exp(-1.1278) ≈ 0.324 – each 1% increase in APR multiplies the odds of acceptance by 0.324 (a 67.6% drop).

##  odd ratio and pertenage change
#The odds ratio is also used to compare sensitivity across segments – a larger magnitude (more negative coefficient) means customers are more price‑sensitive.(Include a short table or sentence:Price Sensitivity: The logistic model estimates that each 1 percentage point increase in APR reduces the odds of acceptance by 67.6% (odds ratio = 0.324). This confirms that customers are highly responsive to price, and small adjustments can have large impacts on acceptance.This makes the results tangible and directly supports the recommendation to lower APRs.)
```{r}
odds_ratio <- exp(coef(model1)["Rate"])
odds_ratio
# percentage change
#Formula: Percentage Change = (Odds Ratio - 1) × 100%
percent_change <- (odds_ratio - 1) * 100
cat("Each 1% increase in APR reduces the odds of acceptance by", 
    abs(round(percent_change, 1)), "%\n")
```
For a 1 percentage point increase in APR, the odds of acceptance are multiplied by 0.234. In other words, the odds decrease by about 67.6% (since 1−0.234=0.676).


##Logistic Curve Plot

```{r}
# ------------------------------------------------------------
# logistic regression plot
# ------------------------------------------------------------

# Predict probabilities for a smooth curve
aprs <- seq(min(segment$Rate), max(segment$Rate), length.out = 200)
pred_data <- data.frame(Rate = aprs)
pred_data$prob <- predict(model1, newdata = pred_data, type = "response")

# Base plot
p <- ggplot() +
  # Raw data points (jittered to show density)
  geom_jitter(data = segment, 
              aes(x = Rate, y = as.numeric(as.character(Outcome))),
              height = 0.02, width = 0, alpha = 0.2, size = 0.8, color = "gray30") +
  # Logistic regression curve
  geom_line(data = pred_data, aes(x = Rate, y = prob), 
            color = "red", size = 1.2) +
  # Optional confidence band (if you want to add)
  # You can add a confidence interval using predict(..., se.fit = TRUE)
  # For simplicity, we skip it here, but can be added if desired.
  labs(
    title = "Logistic Regression: Probability of Acceptance",
    subtitle = paste("Segment: FICO 684–712, Used, 60mo, $17.8K–$25K (n =", nrow(segment), ")"),
    x = "APR (%)", y = "Probability of Acceptance"
  ) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "gray40"),
    panel.grid.minor = element_blank()
  )

print(p)

```
What it does: Plots the raw binary data (jittered) and overlays the fitted S‑curve from the logistic model.

Interpretation: The curve stays between 0 and 1, respecting probability bounds. It clearly shows how acceptance probability declines as APR rises – a key visual for the “Eureka!” moment.

#Expected Profit & Optimal APR

```{r}
# ----------------------------------------------------
# 7. Expected profit per quote and optimal APR
# ----------------------------------------------------
avg_cost <- mean(segment$cost_of_funds, na.rm = TRUE)
avg_amount <- mean(segment$Amount, na.rm = TRUE)
term_years <- 5   # 60 months

# Use a grid with 0.25% steps to include 4.75% if it's the optimum
aprs_grid <- seq(3, 8, by = 0.25)
pred_probs_grid <- predict(model1, newdata = data.frame(Rate = aprs_grid), 
                           type = "response")

profit_df <- data.frame(
  APR = aprs_grid,
  prob = pred_probs_grid,
  profit_if_accept = (aprs_grid - avg_cost) * avg_amount/100 * term_years,
  expected_profit = pred_probs_grid * (aprs_grid - avg_cost) * avg_amount/100 * term_years
)

# Find optimal APR
optimal <- profit_df[which.max(profit_df$expected_profit), ]
cat("\nOptimal APR:", optimal$APR, "%\n")
cat("Max expected profit per quote: $", round(optimal$expected_profit, 2), "\n")

# Current practice
current_apr <- mean(segment$Rate, na.rm = TRUE)
current_prob <- mean(as.numeric(as.character(segment$Outcome)), na.rm = TRUE) # current accept rate
current_profit <- current_prob * (current_apr - avg_cost) * avg_amount/100 * term_years

cat("\nCurrent average APR:", round(current_apr, 2), "%\n")
cat("Current acceptance rate:", round(current_prob * 100, 1), "%\n")
cat("Current expected profit per quote: $", round(current_profit, 2), "\n")
cat("\nImprovement per quote: $", round(optimal$expected_profit - current_profit, 2),
    "(", round((optimal$expected_profit/current_profit - 1)*100, 1), "%)\n")
# Acceptance rate for a specific APR
predict(model1, newdata = data.frame(Rate = 4.75), type = "response")
    
```
What it does: Computes the expected profit per quote for each possible APR, using the model’s predicted probabilities and real business inputs (cost of funds, average loan amount).

##Interpretation:
The optimal APR is the one that maximises expected profit, balancing higher margin vs. lower acceptance.
The optimal APR is 4.5% (or 4.75% depending on the grid). This is the quantitative recommendation to e‑Car.



##

```{r}
# ----------------------------------------------------
# 8. Expected profit plot
# ----------------------------------------------------

ggplot(profit_df, aes(x = APR, y = expected_profit)) +
  geom_line(size = 1, color = "darkgreen") +
  geom_point(data = optimal, aes(x = APR, y = expected_profit), 
             color = "red", size = 3) +
  geom_vline(xintercept = optimal$APR, linetype = "dashed", color = "red") +
  geom_vline(xintercept = current_apr, linetype = "dotted", color = "blue") +
  labs(
    title = "Expected Profit per Quote vs APR",
    x = "APR (%)", y = "Expected Profit ($)"
  ) +
  theme_minimal()
```
## Compute the required quantities for any APR

```{r}
# Example using the segment's actual averages
loan_amount <- mean(segment$Amount)      # e.g., 21847
term_years <- 5                           # 60 months
cost_rate <- mean(segment$cost_of_funds)  # e.g., 1.84%

aprs <- c(6.45, 4.75)   # in percent
probs <- predict(model1, newdata = data.frame(Rate = aprs), type = "response")

revenue <- (aprs / 100) * loan_amount * term_years
cost    <- (cost_rate / 100) * loan_amount * term_years
net     <- revenue - cost
expected <- net * probs

result <- data.frame(
  APR = aprs,
  Revenue = round(revenue),
  Cost = round(cost),
  Net = round(net),
  P_Accept = round(probs, 2),
  Exp_Profit = round(expected, 2)
)
print(result)
```
# Generate the logistic equation text
```{r}
intercept <- coef(model1)[1]
slope <- coef(model1)[2]
cat("Prob(accept) = exp(", round(intercept,2), " + ", round(slope,2), " * APR) / (1 + exp(", round(intercept,2), " + ", round(slope,2), " * APR))", sep="")
```




## Below This graph visually reinforces the teaching note’s point: changing customer characteristics (or market conditions) alters both the baseline acceptance and price sensitivity, which is why personalised pricing matters.

```{r}

# Define APR range
aprs <- seq(4, 10, length.out = 200)

# Create data for varying intercept (a) with fixed slope (b)
a_values <- c(8, 9, 10)   # increasing baseline attractiveness
b_fixed <- -1.2            # fixed slope (negative)

df_a <- expand.grid(APR = aprs, a = a_values)
df_a$b <- b_fixed
df_a$prob <- 1 / (1 + exp(-(df_a$a + df_a$b * df_a$APR)))
df_a$type <- paste0("a = ", df_a$a)

# Create data for varying slope (b) with fixed intercept (a_fixed)
b_values <- c(-1.5, -1.2, -0.9)   # decreasing magnitude (less negative)
a_fixed <- 9

df_b <- expand.grid(APR = aprs, b = b_values)
df_b$a <- a_fixed
df_b$prob <- 1 / (1 + exp(-(df_b$a + df_b$b * df_b$APR)))
df_b$type <- paste0("b = ", df_b$b)

# Combine both datasets
df_plot <- rbind(
  cbind(df_a, set = "Increasing constant (a)"),
  cbind(df_b, set = "Decreasing magnitude of b")
)

# Plot
ggplot(df_plot, aes(x = APR, y = prob, color = type, linetype = set)) +
  geom_line(size = 1.2) +
  labs(
    title = "Logistic Regression: Model Flexibility",
    x = "Quoted APR (%)",
    y = "Probability of Acceptance",
    color = "Parameter",
    linetype = "Effect"
  ) +
  scale_x_continuous(breaks = 4:10) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    legend.box = "vertical",
    plot.title = element_text(face = "bold")
  ) +
  guides(color = guide_legend(nrow = 2))
```
# Explanation of the code:

APR range – set from 4% to 10% (adjust as needed).

Varying intercept (a) – keep slope constant (b = -1.2) and try three increasing values of a. Higher a shifts the curve to the right.

Varying slope (b) – keep intercept constant (a = 9) and try three values of b that are less negative (e.g., -1.5 → -0.9). A less negative slope makes the curve flatter (customers less sensitive) and also shifts it right because the same APR yields higher probability.




```{r}
# ----------------------------------------------------
# 9. Optional: ROC curve (model validation)
# ----------------------------------------------------
library(ROCR)
pred_prob <- predict(model1, type = "response")
pred <- prediction(pred_prob, segment$Outcome)
perf <- performance(pred, "tpr", "fpr")
auc <- performance(pred, "auc")@y.values[[1]]

plot(perf, col = "blue", lwd = 2,
     main = paste("ROC Curve (AUC =", round(auc, 3), ")"))
abline(a = 0, b = 1, lty = 2, col = "gray")
grid()

cat("\nAUC:", round(auc, 3), "\n")

```
What it does: Plots the ROC curve and calculates the Area Under the Curve (AUC).

Interpretation:  Validates model performance (AUC > 0.8 is strong). – the model has excellent discriminatory power. This gives confidence that the predicted probabilities are reliable.


#optimal APR

```{r}
# Calculate optimal APR (like in teaching note)
aprs <- seq(3, 8, by = 0.25)

results <- data.frame(
  APR = aprs,
  prob_accept = predict(model1, 
                        newdata = data.frame(Rate = aprs),
                        type = "response")
)

# Add profit calculation (using actual segment averages, not assumptions)
avg_cost <- mean(segment$cost_of_funds, na.rm = TRUE)
avg_amount <- mean(segment$Amount, na.rm = TRUE)

results <- results %>%
  mutate(
    profit_if_accept = (APR - avg_cost) * avg_amount/100 * (60/12),
    expected_profit = prob_accept * profit_if_accept
  )
# 4. Find best
best <- results[which.max(results$expected_profit), ]
best_apr <- best$APR
best_profit <- best$expected_profit
best_prob <- best$prob_accept

# 5. Compare to current
current_apr <- mean(segment$Rate, na.rm = TRUE)
current_prob <- mean(as.numeric(as.character(segment$Outcome)), na.rm = TRUE)
current_profit <- current_prob * (current_apr - avg_cost) * avg_amount/100 * (60/12)

improvement <- (best_profit - current_profit) / current_profit * 100

print(paste("Optimal APR:", best_apr, "%"))
print(paste("Current avg APR:", current_apr, "%"))

```
# visualization on the profit plot
```{r}
plot(results$APR, results$expected_profit, 
     type = "l", lwd = 3, col = "blue",
     xlab = "APR (%)", 
     ylab = "Expected Profit per Quote ($)",
     main = "Optimal APR Analysis ",
     ylim = c(0, max(results$expected_profit) * 1.1))

abline(v = best_apr, col = "green", lwd = 2, lty = 2)
abline(v = current_apr, col = "red", lwd = 2, lty = 2)
points(best_apr, best_profit, col = "green", pch = 19, cex = 1.5)
points(current_apr, current_profit, col = "red", pch = 19, cex = 1.5)

legend("topright", 
       legend = c("Expected Profit per Quote",
                  paste("Optimal:", round(best_apr, 2), "%"),
                  paste("Current:", round(current_apr, 2), "%")),
       col = c("blue", "green", "red"), 
       lty = c(1, 2, 2), 
       lwd = c(3, 2, 2),
       pch = c(NA, 19, 19),
       pt.cex = 1.5,
       bg = "white")
grid()
```
###Comparison Table and Summary(Creates comparison and prints results for the broader segment.)
```{r}
# Create comparison table for presentation
comparison <- data.frame(
  Metric = c("APR (%)", "Acceptance Rate (%)", "Expected Profit per Loan ($)"),
  Current = c(round(current_apr, 2), 
              round(current_prob * 100, 1),
              paste("$", format(round(current_profit, 2), big.mark = ","))),
  Optimal = c(round(best_apr, 2), 
              round(best$prob_accept * 100, 1),
              paste("$", format(round(best_profit, 2), big.mark = ","))),
  Improvement = c(paste(round(best_apr - current_apr, 2), "pts"),
                  paste("+", round((best$prob_accept - current_prob) * 100, 1), "pts"),
                  paste("+$", format(round(best_profit - current_profit, 2), big.mark = ",")))
)

print(comparison)

# You could also calculate total segment opportunity
total_opportunity <- (best_profit - current_profit) * nrow(segment)
cat("\nTotal opportunity for this segment: $", 
    format(round(total_opportunity), big.mark = ","), "\n")
```
# Results

```{r}
#  Print results
cat("\n", rep("=", 50), "\n", sep = "")
cat(" SEGMENT RESULTS\n")
cat(rep("=", 50), "\n\n", sep = "")

cat(" Segment size:", nrow(segment), "loans\n")
cat(" Avg loan amount: $", format(round(avg_amount), big.mark = ","), "\n")
cat(" Avg cost of funds:", round(avg_cost, 2), "%\n\n")

cat(" CURRENT PRACTICE:\n")
cat("   - Average APR:", round(current_apr, 2), "%\n")
cat("   - Acceptance rate:", round(current_prob * 100, 1), "%\n")
cat("   - Profit if accepted: $", format(round((current_apr - avg_cost) * avg_amount/100 * 5), big.mark = ","), "\n")
cat("   - Expected profit per QUOTE: $", format(round(current_profit, 2), big.mark = ","), "\n\n")

cat(" OPTIMAL PRICING:\n")
cat("   - Recommended APR:", round(best_apr, 2), "%\n")
cat("   - Projected acceptance rate:", round(best_prob * 100, 1), "%\n")
cat("   - Profit if accepted: $", format(round((best_apr - avg_cost) * avg_amount/100 * 5), big.mark = ","), "\n")
cat("   - Expected profit per QUOTE: $", format(round(best_profit, 2), big.mark = ","), "\n\n")

cat(" IMPROVEMENT:\n")
cat("   - Per quote: $", format(round(best_profit - current_profit, 2), big.mark = ","), "\n")
cat("   - Percent: ", round((best_profit - current_profit) / current_profit * 100, 1), "%\n")
cat("   - Segment total: $", format(round((best_profit - current_profit) * nrow(segment)), big.mark = ","), "\n")
```


# example for simple model(just play around the number , not necessary)
```{r}
# ====================================================
# OPTIMAL APR CALCULATION - SIMPLE MODEL (RATE ONLY)
# Using actual model coefficients from segment
# ====================================================

# Use the actual model and segment data
intercept <- coef(model1)[1]      # Should be ~6.36
rate_coef <- coef(model1)[2]      # Should be ~ -1.13

# Segment averages from the actual data
avg_cost <- mean(segment$cost_of_funds, na.rm = TRUE)
avg_amount <- mean(segment$Amount, na.rm = TRUE)
term_years <- 5  # 60 months = 5 years
segment_size <- nrow(segment)  # Should be 1540

# Test APRs from 3% to 8% in 0.25% increments
aprs <- seq(3, 8, by = 0.25)

# Calculate probability of acceptance for each APR
linear_predictor <- intercept + rate_coef * aprs
prob_accept <- 1 / (1 + exp(-linear_predictor))

# Calculate profit if customer accepts
profit_if_accept <- (aprs - avg_cost) * avg_amount/100 * term_years

# Calculate expected profit per quote
expected_profit <- prob_accept * profit_if_accept

# Create results dataframe
results <- data.frame(
  APR = aprs,
  Probability = round(prob_accept, 3),
  Profit_if_Accept = round(profit_if_accept, 0),
  Expected_Profit = round(expected_profit, 0)
)

# Display all results
print("=== EXPECTED PROFIT BY APR ===")
print(results)

# Find optimal APR
best_idx <- which.max(expected_profit)
best_apr <- aprs[best_idx]
best_prob <- prob_accept[best_idx]
best_profit <- expected_profit[best_idx]

# Current segment averages - DYNAMIC from segment
current_apr <- mean(segment$Rate, na.rm = TRUE)
current_prob <- mean(as.numeric(as.character(segment$Outcome)), na.rm = TRUE)
current_profit <- current_prob * (current_apr - avg_cost) * avg_amount/100 * term_years

# Print results
cat("\n", rep("=", 50), "\n", sep = "")
cat("           OPTIMAL APR ANALYSIS\n")
cat(rep("=", 50), "\n\n", sep = "")

cat(" Segment Information:\n")
cat("   Average Loan Amount: $", format(round(avg_amount), big.mark = ","), "\n")
cat("   Average Cost of Funds:", round(avg_cost, 2), "%\n")
cat("   Loan Term: 60 months (5 years)\n")
cat("   Segment Size:", format(segment_size, big.mark = ","), "loans\n\n")

cat(" CURRENT PRACTICE:\n")
cat("   APR:", round(current_apr, 2), "%\n")
cat("   Acceptance Rate:", round(current_prob * 100, 1), "%\n")
cat("   Expected Profit per Quote: $", format(round(current_profit), big.mark = ","), "\n\n")

cat(" OPTIMAL PRICING:\n")
cat("   Recommended APR:", best_apr, "%\n")
cat("   Projected Acceptance Rate:", round(best_prob * 100, 1), "%\n")
cat("   Expected Profit per Quote: $", format(round(best_profit), big.mark = ","), "\n\n")

cat(" IMPROVEMENT:\n")
cat("   APR Change:", round(best_apr - current_apr, 2), "percentage points\n")
cat("   Acceptance Rate Change: +", round((best_prob - current_prob) * 100, 1), "percentage points\n")
cat("   Profit Increase: $", format(round(best_profit - current_profit), big.mark = ","), "per quote\n")
cat("   Percent Improvement:", round((best_profit/current_profit - 1) * 100, 1), "%\n\n")

cat(" Total Segment Opportunity:\n")
cat("   Segment Size:", format(segment_size, big.mark = ","), "loans\n")
cat("   Total Additional Profit: $", 
    format(round((best_profit - current_profit) * segment_size), big.mark = ","), "\n")
```

```{r}

# Create comparison table
comparison <- data.frame(
  Metric = c("APR (%)", 
             "Acceptance Rate (%)", 
             "Profit if Accepted ($)",
             "Expected Profit per Quote ($)",
             "Segment Total Profit ($)"),
  Current = c(round(current_apr, 2),
              round(current_prob * 100, 1),
              format(round((current_apr - avg_cost) * avg_amount/100 * 5), big.mark = ","),
              paste("$", format(round(current_profit), big.mark = ",")),
              paste("$", format(round(current_profit * 2332), big.mark = ","))),
  Optimal = c(round(best_apr, 2),
              round(best_prob * 100, 1),
              format(round((best_apr - avg_cost) * avg_amount/100 * 5), big.mark = ","),
              paste("$", format(round(best_profit), big.mark = ",")),
              paste("$", format(round(best_profit * 2332), big.mark = ","))),
  Change = c(paste(round(best_apr - current_apr, 2), "pts"),
             paste("+", round((best_prob - current_prob) * 100, 1), "pts"),
             paste("$", format(round((best_apr - current_apr) * avg_amount/100 * 5), big.mark = ",")),
             paste("+$", format(round(best_profit - current_profit), big.mark = ",")),
             paste("+$", format(round((best_profit - current_profit) * 2332), big.mark = ",")))
)

library(knitr)
kable(comparison, caption = "Segment: Current vs Optimal Pricing", format = "pipe")
```


```{r}
# ====================================================
# QUICK SCAN OF OTHER SEGMENTS 
# ====================================================

quick_check <- function(data, fico_min, fico_max, car_type_val, term_val, 
                        amount_min = NULL, amount_max = NULL, segment_name = NULL) {
  
  # Filter the data
  seg <- data %>%
    filter(FICO >= fico_min, FICO <= fico_max,
           car_type == car_type_val,
           Term == term_val)
  
  # Apply amount filters if provided
  if (!is.null(amount_min) & !is.null(amount_max)) {
    seg <- seg %>% filter(Amount >= amount_min, Amount <= amount_max)
  }
  
  # Need at least 100 observations for reliable estimates
  if (nrow(seg) < 100) return(NULL)
  
  # Fit logistic model
  m <- glm(Outcome ~ Rate, data = seg, family = binomial)
  
  # Test APRs with finer grid (0.25% steps to match main analysis)
  aprs <- seq(3, 12, by = 0.25)
  probs <- predict(m, newdata = data.frame(Rate = aprs), type = "response")
  
  # Get segment-specific averages
  avg_cost <- mean(seg$cost_of_funds, na.rm = TRUE)
  avg_amount <- mean(seg$Amount, na.rm = TRUE)
  term_years <- seg$Term[1] / 12  # Convert months to years
  
  # Calculate profits
  profit_if_accept <- (aprs - avg_cost) * avg_amount / 100 * term_years
  expected_profit <- probs * profit_if_accept
  
  # Find optimal
  best_idx <- which.max(expected_profit)
  optimal_apr <- aprs[best_idx]
  optimal_prob <- probs[best_idx]
  optimal_profit <- expected_profit[best_idx]
  
  # Current segment metrics
  current_apr <- mean(seg$Rate, na.rm = TRUE)
  current_prob <- mean(as.numeric(as.character(seg$Outcome)), na.rm = TRUE)
  current_profit <- current_prob * (current_apr - avg_cost) * avg_amount / 100 * term_years
  
  # Return results
  data.frame(
    Segment = ifelse(is.null(segment_name),
                     paste(fico_min, "-", fico_max, car_type_val, term_val, "mo"),
                     segment_name),
    N = nrow(seg),
    Current_APR = round(current_apr, 2),
    Optimal_APR = round(optimal_apr, 2),
    APR_Gap = round(optimal_apr - current_apr, 2),
    Current_Profit = paste0("$", format(round(current_profit), big.mark = ",")),
    Optimal_Profit = paste0("$", format(round(optimal_profit), big.mark = ",")),
    Improvement = paste0(round((optimal_profit / current_profit - 1) * 100, 1), "%"),
    stringsAsFactors = FALSE
  )
}

# Run the scan on selected segments
segment_results <- bind_rows(
  quick_check(df, 684, 712, "U", 60, 17800, 25000, "Teaching Segment"),
  quick_check(df, 712, 780, "N", 60, 20000, 40000, "Prime New Cars"),
  quick_check(df, 620, 680, "U", 48, 10000, 20000, "Near-prime Used"),
  quick_check(df, 580, 620, "U", 72, 15000, 25000, "Subprime Used"),
  quick_check(df, 700, 750, "R", 60, 15000, 30000, "Prime Refinance"),
  quick_check(df, 760, 850, "N", 60, 30000, 60000, "Super-prime New"),
  quick_check(df, 650, 700, "U", 60, 15000, 25000, "Mid-range Used"),
  quick_check(df, 650, 720, "U", 36, 10000, 20000, "Short-term Used")
)

# Remove any NULL results
segment_results <- segment_results[!sapply(segment_results, is.null)]
segment_results <- bind_rows(segment_results)

# Display results with nice formatting
library(knitr)
library(kableExtra)

segment_results %>%
  kable(format = "html", 
        caption = "<b>Table: Profit Optimization Across Segments</b>",
        align = c("l", "r", "r", "r", "r", "r", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), 
                full_width = FALSE,
                position = "left") %>%
  row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
  row_spec(which(segment_results$Segment == "Teaching Segment"), 
           bold = TRUE, background = "#e8f4f8")

# Summary insights
cat("\n", rep("=", 80), "\n", sep = "")
cat("KEY FINDINGS FROM SEGMENT SCAN:\n")
cat(rep("=", 80), "\n\n", sep = "")

overpriced <- sum(segment_results$APR_Gap < 0, na.rm = TRUE)
cat(" •", overpriced, "out of", nrow(segment_results), 
    "segments show overpricing (optimal APR < current APR)\n")

avg_improvement <- mean(as.numeric(gsub("%", "", segment_results$Improvement)), na.rm = TRUE)
cat(" • Average improvement potential:", round(avg_improvement, 1), "%\n")

best_idx <- which.max(as.numeric(gsub("%", "", segment_results$Improvement)))
cat(" • Biggest opportunity:", segment_results$Segment[best_idx], 
    "(", segment_results$Improvement[best_idx], "improvement)\n")

# Visualize APR gaps
if (nrow(segment_results) > 0) {
  # Create a simple barplot
  barplot(segment_results$APR_Gap,
          names.arg = substr(segment_results$Segment, 1, 15),
          las = 2,
          col = ifelse(segment_results$APR_Gap < 0, "lightcoral", "lightgreen"),
          main = "APR Gap by Segment (Optimal - Current)",
          xlab = "Segment", ylab = "APR Gap (percentage points)",
          cex.names = 0.8)
  abline(h = 0, col = "black", lwd = 2)
  grid()
}
```

#This shows how to scale – you might keep one as a teaser for “next steps”.
# Model with multivariables(Fits and interprets models with Rate + FICO + Amount + Term, and finds optimal APR holding other variables constant.)
```{r}
# These are advanced analyses that go beyond the teaching note. 
# Model with Rate, FICO, Amount, and Term
model_multiple <- glm(Outcome ~ Rate + FICO + Amount + Term, 
                      family = binomial(link = "logit"),
                      data = segment )

summary(model_multiple)

```
## Interpret the Output
```{r}
# Look at coefficients
coef_summary <- summary(model_multiple)$coefficients
print(coef_summary)

# Calculate odds ratios
odds_ratios <- exp(coef(model_multiple))
print(round(odds_ratios, 3))

```

# example to calculate

```{r}
# Use your actual model
model_multiple <- glm(Outcome ~ Rate + FICO + Amount, 
                      family = binomial(link = "logit"),
                      data = segment)

# Create typical customer
typical <- data.frame(
  Rate = seq(3, 8, by = 0.25),
  FICO = 702,
  Amount = 21847
)

# Predict probabilities
typical$prob <- predict(model_multiple, newdata = typical, type = "response")

# Calculate profit
avg_cost <- mean(segment$cost_of_funds, na.rm = TRUE)
typical$profit_if_accept <- (typical$Rate - avg_cost) * 21847/100 * 5
typical$expected_profit <- typical$prob * typical$profit_if_accept

# Find optimal
best <- typical[which.max(typical$expected_profit), ]
print(paste("Optimal APR:", best$Rate, "%"))
print(paste("Expected Profit: $", round(best$expected_profit, 2)))
```

# Interaction check,These show that price sensitivity varies by term, FICO, etc. – supporting the move to individual pricing. Keep if you want to demonstrate the broader value of Nomis’s approach.

```{r}
# ====================================================
# INTERACTION TESTS ON FULL DATASET
# ====================================================

# 1. Rate × Term Interaction (Does price sensitivity vary by loan term?)
model_term_interaction <- glm(Outcome ~ Rate * Term + FICO + Amount + car_type, 
                              family = binomial(link = "logit"), 
                              data = df)

summary(model_term_interaction)
# Look for: Rate:Term coefficient and p-value

# 2. Rate × FICO Interaction (Does price sensitivity vary by credit score?)
model_fico_interaction <- glm(Outcome ~ Rate * FICO + Term + Amount + car_type, 
                              family = binomial(link = "logit"), 
                              data = df)

summary(model_fico_interaction)
# Look for: Rate:FICO coefficient and p-value

# 3. Rate × Amount Interaction (Does price sensitivity vary by loan size?)
model_amount_interaction <- glm(Outcome ~ Rate * Amount + FICO + Term + car_type, 
                                family = binomial(link = "logit"), 
                                data = df)

summary(model_amount_interaction)
# Look for: Rate:Amount coefficient and p-value

# 4. Rate × Car Type Interaction (Does price sensitivity vary by new/used/refi?)
model_cartype_interaction <- glm(Outcome ~ Rate * car_type + FICO + Term + Amount, 
                                 family = binomial(link = "logit"), 
                                 data = df)

summary(model_cartype_interaction)
# Look for: Rate:car_type coefficients

## What it does: Tests whether price sensitivity varies with term, FICO, amount, or car type.

##Interpretation: All interactions are highly significant (p < 0.001), proving that one APR does not fit all. This justifies moving from segment‑level to individual‑level pricing – the ultimate scalable solution Nomis offers.
```
```{r}
# Extract interaction p-values
library(knitr)

# Run models
m_term <- glm(Outcome ~ Rate * Term + FICO + Amount + car_type, 
              family = binomial, data = df)
m_fico <- glm(Outcome ~ Rate * FICO + Term + Amount + car_type, 
              family = binomial, data = df)
m_amount <- glm(Outcome ~ Rate * Amount + FICO + Term + car_type, 
                family = binomial, data = df)
m_cartype <- glm(Outcome ~ Rate * car_type + FICO + Term + Amount, 
                 family = binomial, data = df)

# Create results table
interaction_results <- data.frame(
  Interaction = c("Rate × Term", "Rate × FICO", "Rate × Amount", "Rate × Car Type"),
  Coefficient = c(
    coef(m_term)["Rate:Term"],
    coef(m_fico)["Rate:FICO"],
    coef(m_amount)["Rate:Amount"],
    "Multiple coefficients"  # Car type has multiple levels
  ),
  P_value = c(
    summary(m_term)$coefficients["Rate:Term", 4],
    summary(m_fico)$coefficients["Rate:FICO", 4],
    summary(m_amount)$coefficients["Rate:Amount", 4],
    "See below"
  ),
  Significant = c(
    ifelse(summary(m_term)$coefficients["Rate:Term", 4] < 0.05, "YES", "NO"),
    ifelse(summary(m_fico)$coefficients["Rate:FICO", 4] < 0.05, "YES", "NO"),
    ifelse(summary(m_amount)$coefficients["Rate:Amount", 4] < 0.05, "YES", "NO"),
    "Check car_type interactions"
  )
)

print(interaction_results)

# For car type, show individual interactions
cat("\n--- Rate × Car Type Interactions ---\n")
summary(m_cartype)$coefficients[grep("Rate:car_type", rownames(summary(m_cartype)$coefficients)), ]
```



